---
title: 'Statistical Inference Part 1: Simulation Exercise'
author: "JackHo"
date: "30 October 2016"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Before Cooking

*To keep the article is shorter than 3 pages, some codes have been shaded by setting `echo=FALSE' in the chunk option.*

```{r Loadinglibs, warning= FALSE, message= FALSE}
# Loading necessary packages
require(ggplot2)
require(dplyr)
```

## Overview
**In this project I investigated the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. I investigated the distribution of averages of 40 exponentials.To get better illsutrations, I simulated the expoential distribution `100000` times.**

## Part 1: Simulation Exercise Instructions
### 1.1 Compare the sample mean with the theoretical mean of expoential distribution
```{r Exponentials_mean, cache=TRUE}
# Set seed- keeping its reproducible
set.seed(1)
# Sim 100000 times and calculate means
SMeanRexp <- vector(length = 100000, mode = "numeric")
CMeanRexp <- vector(length = 100000, mode = "numeric")
SVarianceRexp <- vector(length = 100000, mode = "numeric")
CVarianceRexp <- vector(length = 100000, mode = "numeric")
vec <- vector(length = 40, mode = "numeric")
for(i in 1:100000){
      vec<-rexp(n = 40, rate = .2);      SMeanRexp[i] <- mean(vec)
      SVarianceRexp[i] <- var(vec);      CMeanRexp[i] <- sum(SMeanRexp)/i
      CVarianceRexp[i] <- sum(SVarianceRexp)/i
}
```
```{r hiddencodes1, echo=FALSE}
SMeanRexp <- data.frame(SampleMean = SMeanRexp)
SRexp <- SMeanRexp
SRexp$TheoreMean <- rep(1/.2,100000)
SRexp$CumMean <- CMeanRexp
SRexp$SampleVariance <- SVarianceRexp
SRexp$CumVariance <- CVarianceRexp
```
```{r cache=TRUE}
print(sprintf("The mean of the sample is:  %f The Theoretical mean is: %f", mean(SRexp[,1]), 1/.2))
```

The `sample mean` (`r mean(SRexp[,1])`) of the distribution is a little larger than the `theoretical mean` (`r 1/.2 `) of the distribution.

### 1.2 Compare the sample variance with the theoretical variance of expoential distribution
```{r Exponentials_variance, cache=TRUE}
print(sprintf("The mean of the variance is:  %f The Theoretical variance is: %f", mean(SRexp$SampleVariance),(1/.2)^2))
```

The `sample variance` (`r mean(SRexp$SampleVariance)`) of the distribution is a little larger than the `theoretical mean` (`r (1/.2)^2 `) of the distribution. 

### 1.3 Visulaze the distributions and trends
```{r Exponentials_mean_plots1, fig.align = 'center', cache = TRUE , fig.height=5, fig.width=6}
# Draw histogram- checking the distribution of 1000 times exponential distribution
ggplot(SRexp,aes(SampleMean)) + geom_histogram(fill="salmon",aes(y= ..density..),binwidth = .05)+stat_function(fun = dnorm, color = "brown", size =1.1, linetype =4,args = list(mean=5))+ggtitle("THe Histogram of 100000 Expoential Distribution Simulations")+ geom_vline(xintercept = c(5,mean(SRexp$SampleMean)),size=.5,col="black",linetype=2)+theme_bw() # the difference is too small to distinguish the distance between two vertical lines
```

From the plot `THe Histogram of the simulation of 100000 Expoential Distribution`, the distribution is highly close to Gaussian (very look like a bell-shape). 

```{r Exponentials_mean_plots2, fig.align = 'center', cache = TRUE , fig.height=5, fig.width=6}
# Plot the trends of the cumulated sample means
ggplot(SRexp,aes(y=CMeanRexp,x=1:100000))+geom_line(col="salmon",size = 1.5)+geom_hline(yintercept = 1/.2,col="#9cb6ff",linetype=2.2,size=1.2) + xlab("The Number of Experiments") + ylab("Cumulated Means Samples")+ggtitle(label = "The Cumulated Means of 100000 Samples", subtitle ="compared with the Theoretical Mean")+theme_bw()
```

From the plot `The Cumulated Mean of 100000 Expoential Distribution compared with the ` `Theoretical Mean of 100000 Expoential Distribution`, the cumulated tends to be close to the theoretical mean after around 25000 times experiments.   
*The codes of plotting the graph below have been shaded.*
```{r Exponentials_variance_plots, fig.align = 'center', cache = TRUE, echo = FALSE , fig.height=5, fig.width=6}
# Plot the trends of the cumulated variance
ggplot(SRexp,aes(y=CumVariance,x=1:100000))+geom_line(col="salmon",size = 1.5)+geom_hline(yintercept = (1/.2)^2,col="#9cb6ff",linetype=2,size=1.1) + xlab("The Number of Experiments") + ylab("The Means of the Cumulated Variance of Experiments")+ggtitle(label = "The Means of the Cumulated Variance of Experiments", subtitle ="compared with the Theoretical Variance ")+theme_bw()
```

Bascially, like the things between the mean of the sample means, the variance of the sample will ultimately be close to the population variance. See the plot below.

### 1.4 Summary 
As a whole, from the 2 sections above, based on a large number of experiments, the mean of the sample means will be gravitated toward to the population mean. As a result, we could concluded that the mean of the samples can predict the mean of the population, and the variance of the sample, can also predict the variance of the population variance.
